### 2009-07-17 paper.tex
converted figs to EPS
fixed CYK bug (bifurcations were using sum instead of max)

% figures in figs_plos:
%  1: inside.pdf
%  2: insidetransemit.pdf
%  3: cyk.pdf
%  4: cyktransemit.pdf
%  5: cyktraceback.pdf
%  6: outside.pdf
%  7: outsidetransemit.pdf
%  8: stree-with-structure.pdf
%  9: stree-mutations-with-structures.pdf
% 10: perfect.pdf
% 11: ama.pdf

@@

\begin{figure}[!ht]
% 
%   \SetKwData{maxProb}{maxProb}
%   \SetKwData{bifurcProb}{bifurcProb}
%   \SetKwFunction{calcTransEmitProb}{calcTransEmitProb}
%   \SetKwFunction{calcLBifurcProb}{calcLBifurcProb}
%   \SetKwFunction{calcRBifurcProb}{calcRBifurcProb}
%   \SetLine
%   \KwIn{sequences $X, Y, Z$}
% 
%   \ForEach(\tcc*[f]{inside$\to$outside sorted}){$n^{(X)} \in \foldenv^{(X)}$} {
%     \ForEach(\tcc*[f]{inside$\to$outside sorted}){$n^{(Y)} \in \foldenv^{(Y)}$} {
%       \ForEach(\tcc*[f]{inside$\to$outside sorted}){$n^{(Z)} \in \foldenv^{(Z)}$} {
%         \BlankLine
%         \ForEach{state $\bvec{a}$}{
%           \BlankLine
%           \bifurcProb $\leftarrow 0$\;
%           \ForEach{$\left( n^{(X)}_L,n^{(X)}_R \right) \in b_{in}\left( n^{(X)} \right)$} {
%             \ForEach{$\left( n^{(Y)}_L,n^{(Y)}_R \right) \in b_{in}\left( n^{(Y)} \right)$} {
%               \ForEach{$\left( n^{(Z)}_L,n^{(Z)}_R \right) \in b_{in}\left( n^{(Z)} \right)$} {
%                 \bifurcProb += \calcLBifurcProb($\bvec{a}; \cdot$)\;
%                 \bifurcProb += \calcRBifurcProb($\bvec{a}; \cdot$)\;
%               }
%             }
%           }
%           \BlankLine
%           $\gamma_{\bvec{a}} \left( n^{(X)}, n^{(Y)}, n^{(Z)} \right)$ \\
%           $\qquad \leftarrow \max \left( \calcTransEmitProb \left( \bvec{a}; n^{(X)}, n^{(Y)}, n^{(Z)} \right), \bifurcProb \right)$ \;
%           store $\gamma_{\bvec{a}} \left( n^{(X)}, n^{(Y)}, n^{(Z)} \right)$\;
%         }
%         \BlankLine
%       }
%     }
%   }
%   \KwRet{$\gamma_{\bvec{a}}\left( n^{(X)}[0,L^{(X)}],n^{(Y)}[0,L^{(Y)}],n^{(Z)}[0,L^{(Z)}] \right)$}\;
%   
  \caption{\alglabel{cyk}
{\bf    The constrained CYK algorithm for three sequences $X, Y, Z$. }
    Ensemble states $\bvec{a}$ in the iteration over states are sorted in Inside fill order
    with $\Semit$ states first, then $\Snull$ states in reverse topological order.
The subroutine $\mathrm{calcTransEmitProb}()$ is given in \algref{cyktransemit}.
The algorithm to find the Maximum Likelihood parse tree by traceback through the CYK matrix is given in \algref{cyktraceback}.
  }
\end{figure}

@@

\begin{figure}[!ht]
% 
%   \SetKw{Select}{select}
%   \SetKw{Pop}{pop}
%   \SetKw{Push}{push}
%   \SetKw{Goto}{goto}
%   \SetKw{Mainloop}{main loop}
%   \SetKwData{coordinateStack}{coordinateStack}
%   \SetKwFunction{goto}{goto}
%   \SetKwBlock{Begin}{begin main loop:}{end}
%   \SetLine
%   \KwIn{sequences $X, Y, Z$}
% 
%   pushdown stack coordinateStack; \tcc*[f]{hold (state, subsequence triplet) pairs} \\
%   $\bvec{a} \leftarrow \Sstart$; \tcc*[f]{current ensemble state} \\
%   $n^{(X)} \leftarrow N^{(X)}$; \tcc*[f]{current $X$ subsequence; $N^{(X)}$ is the outermost subsequence} \\
%   $n^{(Y)} \leftarrow N^{(Y)}$; \tcc*[f]{current $Y$ subsequence; $N^{(Y)}$ is the outermost subsequence} \\
%   $n^{(Z)} \leftarrow N^{(Z)}$; \tcc*[f]{current $Z$ subsequence; $N^{(Z)}$ is the outermost subsequence} \\
%   clear \coordinateStack;\
% 
%   \Begin{
%     output current state $\bvec{a}$ and subsequence triplet $\left( n^{(X)}, n^{(Y)}, n^{(Z)} \right)$\;
%     \SetLine
%     \uIf(\tcc*[f]{end of a parse subtree}){$\bvec{a}$ is the $\Send$ state}{
%       \lIf(\tcc*[f]{end of the parse tree}){\coordinateStack is empty}{\Return}
%       \Pop $\left( \bvec{a}, n^{(X)}, n^{(Y)}, n^{(Z)} \right)$ from \coordinateStack\;
%       \Goto \Mainloop\;
%     }
%     \uElseIf(\tcc*[f]{bifurcation $\bvec{a} \to \bvec{c} \bvec{b}$}){$\bvec{a}$ is a bifurcation state}{
%       \Select $\left( n^{(X)}_L, n^{(X)}_R \right) \in b_{in}\left( n^{(X)} \right)$,
%       $\left( n^{(Y)}_L, n^{(Y)}_R \right) \in b_{in}\left( n^{(Y)} \right)$,
%       $\left( n^{(Z)}_L, n^{(Z)}_R \right) \in b_{in}\left( n^{(Z)} \right)$ \\
%       such that \\
%       $\gamma_{\bvec{a}} \left(n^{(X)}, n^{(Y)}, n^{(Z)} \right) = \gamma_{\bvec{a}} \left(n^{(X)}_L, n^{(Y)}_L, n^{(Z)}_L \right)  \gamma_{\bvec{a}} \left(n^{(X)}_R, n^{(Y)}_R, n^{(Z)}_R \right)$ \;
%       \Push $\left( \bvec{c}, n^{(X)}_R, n^{(Y)}_R, n^{(Z)}_R \right)$ onto \coordinateStack\;
%       $\bvec{a} \leftarrow \bvec{b}$\;
%       $n^{(X)} \leftarrow n^{(X)}_L$\;
%       $n^{(Y)} \leftarrow n^{(Y)}_L$\;
%       $n^{(Z)} \leftarrow n^{(Z)}_L$\;
%       \Goto \Mainloop\;
%     }
%     \Else(\tcc*[f]{$\Semit$ or $\Snull$ state}){
%       $m^{(X)} \leftarrow c_{in}\left(\bvec{b}; n^{(X)}\right)$ \;
%       $m^{(Y)} \leftarrow c_{in}\left(\bvec{b}; n^{(Y)}\right)$ \;
%       $m^{(Z)} \leftarrow c_{in}\left(\bvec{b}; n^{(Z)}\right)$ \;
%       \Select $\bvec{b} \in \left\{ b : \exists \,\bvec{a} \to \bvec{l}\bvec{b}\bvec{r} \right\}$ \\
%       such that \\
%       $\gamma_{\bvec{a}} \left( n^{(X)}, n^{(Y)}, n^{(Z)} \right) = P(\bvec{a} \to \bvec{l}\bvec{b}\bvec{r}) \gamma_{\bvec{b}} \left( m^{(X)}, m^{(Y)}, m^{(Z)} \right)$\;
%       $\bvec{a} \leftarrow \bvec{b}$\;
%       $n^{(X)} \leftarrow m^{(X)}$\;
%       $n^{(Y)} \leftarrow m^{(Y)}$\;
%       $n^{(Z)} \leftarrow m^{(Z)}$\;
%       \Goto \Mainloop\;
%     }
%   }
  \caption{\alglabel{cyktraceback}
{\bf     The constrained CYK traceback algorithm for three sequences $X, Y, Z$.}
The CYK algorithm itself is given in \algref{cyk}.
  }
\end{figure}

>>>>

% figures in figs_plos:
%  1: inside.eps
%  2: insidetransemit.eps
%  3: cyk.eps
%  4: cyktransemit.eps
%  5: cyktraceback.eps
%  6: outside.eps
%  7: outsidetransemit.eps
%  8: stree-with-structure.eps
%  9: stree-mutations-with-structures.eps
% 10: perfect.eps
% 11: ama.eps

@@

\begin{figure}[!ht]
% 
%   \SetKwData{maxProb}{maxProb}
%   \SetKwData{bifurcProb}{bifurcProb}
%   \SetKwFunction{calcTransEmitProb}{calcTransEmitProb}
%   \SetKwFunction{calcLBifurcProb}{calcLBifurcProb}
%   \SetKwFunction{calcRBifurcProb}{calcRBifurcProb}
%   \SetLine
%   \KwIn{sequences $X, Y, Z$}
% 
%   \ForEach(\tcc*[f]{inside$\to$outside sorted}){$n^{(X)} \in \foldenv^{(X)}$} {
%     \ForEach(\tcc*[f]{inside$\to$outside sorted}){$n^{(Y)} \in \foldenv^{(Y)}$} {
%       \ForEach(\tcc*[f]{inside$\to$outside sorted}){$n^{(Z)} \in \foldenv^{(Z)}$} {
%         \BlankLine
%         \ForEach{state $\bvec{a}$}{
%           \BlankLine
%           \bifurcProb $\leftarrow 0$\;
%           \ForEach{$\left( n^{(X)}_L,n^{(X)}_R \right) \in b_{in}\left( n^{(X)} \right)$} {
%             \ForEach{$\left( n^{(Y)}_L,n^{(Y)}_R \right) \in b_{in}\left( n^{(Y)} \right)$} {
%               \ForEach{$\left( n^{(Z)}_L,n^{(Z)}_R \right) \in b_{in}\left( n^{(Z)} \right)$} {
%                 $\bifurcProb \leftarrow \max \left( \bifurcProb, \calcLBifurcProb(\bvec{a}; \cdot) \right)$\;
%                 $\bifurcProb \leftarrow \max \left( \bifurcProb, \calcRBifurcProb(\bvec{a}; \cdot) \right)$\;
%               }
%             }
%           }
%           \BlankLine
%           $\gamma_{\bvec{a}} \left( n^{(X)}, n^{(Y)}, n^{(Z)} \right)$ \\
%           $\qquad \leftarrow \max \left( \calcTransEmitProb \left( \bvec{a}; n^{(X)}, n^{(Y)}, n^{(Z)} \right), \bifurcProb \right)$ \;
%           store $\gamma_{\bvec{a}} \left( n^{(X)}, n^{(Y)}, n^{(Z)} \right)$\;
%         }
%         \BlankLine
%       }
%     }
%   }
%   \KwRet{$\gamma_{\bvec{a}}\left( n^{(X)}[0,L^{(X)}],n^{(Y)}[0,L^{(Y)}],n^{(Z)}[0,L^{(Z)}] \right)$}\;
%   
  \caption{\alglabel{cyk}
{\bf    The constrained CYK algorithm for three sequences $X, Y, Z$. }
    Ensemble states $\bvec{a}$ in the iteration over states are sorted in Inside fill order
    with $\Semit$ states first, then $\Snull$ states in reverse topological order.
The subroutine $\mathrm{calcTransEmitProb}()$ is given in \algref{cyktransemit}.
The algorithm to find the Maximum Likelihood parse tree by traceback through the CYK matrix is given in \algref{cyktraceback}.
  }
\end{figure}

@@

\begin{figure}[!ht]
% 
%   \SetKw{Select}{select}
%   \SetKw{Pop}{pop}
%   \SetKw{Push}{push}
%   \SetKw{Goto}{goto}
%   \SetKw{Mainloop}{main loop}
%   \SetKwData{coordinateStack}{coordinateStack}
%   \SetKwFunction{goto}{goto}
%   \SetKwBlock{Begin}{begin main loop:}{end}
%   \SetLine
%   \KwIn{sequences $X, Y, Z$}
% 
%   pushdown stack coordinateStack; \tcc*[f]{hold (state, subsequence triplet) pairs} \\
%   $\bvec{a} \leftarrow \Sstart$; \tcc*[f]{current ensemble state} \\
%   $n^{(X)} \leftarrow N^{(X)}$; \tcc*[f]{current $X$ subsequence; $N^{(X)}$ is the outermost subsequence} \\
%   $n^{(Y)} \leftarrow N^{(Y)}$; \tcc*[f]{current $Y$ subsequence; $N^{(Y)}$ is the outermost subsequence} \\
%   $n^{(Z)} \leftarrow N^{(Z)}$; \tcc*[f]{current $Z$ subsequence; $N^{(Z)}$ is the outermost subsequence} \\
%   clear \coordinateStack;\
% 
%   \Begin{
%     output current state $\bvec{a}$ and subsequence triplet $\left( n^{(X)}, n^{(Y)}, n^{(Z)} \right)$\;
%     \SetLine
%     \uIf(\tcc*[f]{end of a parse subtree}){$\bvec{a}$ is the $\Send$ state}{
%       \lIf(\tcc*[f]{end of the parse tree}){\coordinateStack is empty}{\Return}
%       \Pop $\left( \bvec{a}, n^{(X)}, n^{(Y)}, n^{(Z)} \right)$ from \coordinateStack\;
%       \Goto \Mainloop\;
%     }
%     \uElseIf(\tcc*[f]{bifurcation $\bvec{a} \to \bvec{c} \bvec{b}$}){$\bvec{a}$ is a bifurcation state}{
%       \Select $\left( n^{(X)}_L, n^{(X)}_R \right) \in b_{in}\left( n^{(X)} \right)$,
%       $\left( n^{(Y)}_L, n^{(Y)}_R \right) \in b_{in}\left( n^{(Y)} \right)$,
%       $\left( n^{(Z)}_L, n^{(Z)}_R \right) \in b_{in}\left( n^{(Z)} \right)$ \\
%       such that \\
%       $\gamma_{\bvec{a}} \left(n^{(X)}, n^{(Y)}, n^{(Z)} \right) = \gamma_{\bvec{a}} \left(n^{(X)}_L, n^{(Y)}_L, n^{(Z)}_L \right)  \gamma_{\bvec{a}} \left(n^{(X)}_R, n^{(Y)}_R, n^{(Z)}_R \right)$ \;
%       \Push $\left( \bvec{c}, n^{(X)}_R, n^{(Y)}_R, n^{(Z)}_R \right)$ onto \coordinateStack\;
%       $\bvec{a} \leftarrow \bvec{b}$\;
%       $n^{(X)} \leftarrow n^{(X)}_L$\;
%       $n^{(Y)} \leftarrow n^{(Y)}_L$\;
%       $n^{(Z)} \leftarrow n^{(Z)}_L$\;
%       \Goto \Mainloop\;
%     }
%     \Else(\tcc*[f]{$\Semit$ or $\Snull$ state}){
%       $m^{(X)} \leftarrow c_{in}\left(\bvec{b}; n^{(X)}\right)$ \;
%       $m^{(Y)} \leftarrow c_{in}\left(\bvec{b}; n^{(Y)}\right)$ \;
%       $m^{(Z)} \leftarrow c_{in}\left(\bvec{b}; n^{(Z)}\right)$ \;
%       \Select $\bvec{b} \in \left\{ b : \exists \,\bvec{a} \to \bvec{l}\bvec{b}\bvec{r} \right\}$ \\
%       such that \\
%       $\gamma_{\bvec{a}} \left( n^{(X)}, n^{(Y)}, n^{(Z)} \right) = P(\bvec{a} \to \bvec{l}\bvec{b}\bvec{r}) \gamma_{\bvec{b}} \left( m^{(X)}, m^{(Y)}, m^{(Z)} \right)$\;
%       $\bvec{a} \leftarrow \bvec{b}$\;
%       $n^{(X)} \leftarrow m^{(X)}$\;
%       $n^{(Y)} \leftarrow m^{(Y)}$\;
%       $n^{(Z)} \leftarrow m^{(Z)}$\;
%       \Goto \Mainloop\;
%     }
%   }
  \caption{\alglabel{cyktraceback}
{\bf     The constrained CYK traceback algorithm for three sequences $X, Y, Z$.}
For compactness we show only the ``if'' case for left-bifurcations ($\bvec{a} \to \bvec{c} \bvec{b}$);
the case for right-bifurcations ($\bvec{a} \to \bvec{b} \bvec{d}$) is identical,
with $(\bvec{c},\bvec{b})$ respectively replaced by $(\bvec{b},\bvec{d})$.
The CYK algorithm itself is given in \algref{cyk}.
  }
\end{figure}
